---
title: "MSMem Multiple Regression Model Comparisons"
output: html_notebook
---

Analyzes variance explained by Matrix Similarity and Mean ATL Degree predictors and compares to clinical and demographic variables. 

V1.0 

Sam Audrain, December 2022
 
# Read in the packages
```{r}

library(car)
library(jtools)

```

# Read in the data
```{r}
###### CHANGE PATH FOR YOUR DATA #######
data<-read.csv("/Users/audrainsp/Library/Mobile Documents/com~apple~CloudDocs/Documents/Science_Projects/RSMemory/MatrixSim_Mem_Nov2020/MSMemNov2020_R/MSMem_Github/R_MSMem.csv")

# subset 
po_data<-subset(data, Surgery == "LTLAmHipp" | Surgery == "RTLAmHipp" ) # excluding controls and selective surgery patients 


```


# Prepare the data 

To statistically test, we will z-score across groups at each threshold, average the z-scored degree across thresholds, and then correlate once

```{r}

# z score across thresholds, for each hemisphere 
for (cost in c("05","10","15","20","25","30","35","40","45")){
  for (hem in c("L","R")){
    eval(parse(text = paste0("po_data$",hem,"_cost",cost,"_z<-scale(po_data$",hem,"_cost",cost,")")))
  }
}

# average across thresholds 
po_data$mean_L_degree<-rowMeans(po_data[ , c("L_cost05_z","L_cost10_z","L_cost15_z","L_cost20_z","L_cost25_z","L_cost30_z","L_cost35_z","L_cost40_z","L_cost45_z")])

po_data$mean_R_degree<-rowMeans(po_data[ , c("R_cost05_z","R_cost10_z","R_cost15_z","R_cost20_z","R_cost25_z","R_cost30_z","R_cost35_z","R_cost40_z","R_cost45_z")])

## center MS
po_data$MatrixSim_64_scaled<-scale(po_data$MatrixSim_64, center=TRUE, scale=FALSE)

```

# LTLE

## Models

### Verbal memory change predicted by matrix similarity + degree
```{r}

# model 
model<-lm(VerbalFactor ~ MatrixSim_64_scaled + mean_L_degree, data=po_data, na.action=na.omit, subset=Group=="LTLE")

# check model assumptions 
plot(model) # residuals vs fitted should be straight, no bad outliers
hist(residuals(model)) # residuals should be normal
shapiro.test(residuals(model)) # residuals should be normal
ncvTest(model) # test for heterogeneity of variance
vif(model) ## variance inflation should be less than 10 

# stats 
summary(model) ##when using just full ATL cases, both degree and MS predicts outcome 

# variance explained 
summ(model, part.corr=TRUE) ### part.r is semipartial corr, partial.r is partial corr. 
#semi-partial correlation = does A influence B above and beyond C influence on B? 
#partial correlation = what is the correlation between A and B after accounting for fact that C influences both? 
## to get unique variance, we want semi-partial corrs. to get shared variance, subtract sub of semi-partial from total model r2 
# remember these are rs, need to square to get % variance explained 
```

### Verbal memory change predicted by matrix similarity + degree + clinical and demographic variables 
```{r}
## model 
model2<-lm(VerbalFactor ~ MatrixSim_64_scaled + mean_L_degree + Pre_Verb + MTS_status +age1 +sex +edu +AOO, data=po_data, na.action=na.omit, subset=Group=="LTLE")

# check model assumptions 
plot(model2) # residuals vs fitted should be straight, no bad outliers
hist(residuals(model2)) # residuals should be normal
shapiro.test(residuals(model2)) # residuals should be normal
ncvTest(model2) # test for heterogeneity of variance
vif(model2) ## variance inflation should be less than 10 

# stats 
summary(model2) ##when using just full ATL cases, both degree and MS predicts outcome 

# variance explained 
summ(model2, part.corr=TRUE) ### part.r is semipartial corr, partial.r is partial corr. 
#semi-partial correlation = does A influence B above and beyond C influence on B? 
#partial correlation = what is the correlation between A and B after accounting for fact that C influences both? 
## to get unique variance, we want semi-partial corrs. to get shared variance, subtract sub of semi-partial from total model r2 
# remember these are rs, need to square to get % variance explained 
```

# Compare models with and without clinical and demographic variables 
```{r}

anova(model,model2)

```

# RTLE

## Models

### Visual memory change predicted by clinical and demographic variables

We already know matrix similarity and degree are not associated with visual memory change, so we don't include them here. 

```{r}
# model 
model<-lm(VisualFactor ~ Pre_Vis +MTS_status +age1 +sex +edu +AOO, data=po_data, na.action=na.omit, subset=Group=="RTLE") 

# check model assumptions 
plot(model2) # residuals vs fitted should be straight, no bad outliers
hist(residuals(model2)) # residuals should be normal
shapiro.test(residuals(model2)) # residuals should be normal
ncvTest(model2) # test for heterogeneity of variance
vif(model2) ## variance inflation should be less than 10 

# stats 
summary(model)



```

