---
title: "MSMem Mean ATL Degree Analysis"
output: html_notebook
---

V2.0 edits based on peer review 

Sam Audrain, September 2023

This script takes already calculated ATL degree and correlates it with material-specific memory change in left and right TLE groups. 


# Read in the packages

```{r}

library(Rmisc)
library(rstatix)
library(ggplot2)
library(dplyr)
library(car)
library(emmeans)

```

# Read in the data 
```{r}
###### CHANGE PATH FOR YOUR DATA #######
data<-read.csv("/Users/samanthaaudrain/Documents/Science_Projects/RSMemory/MatrixSim_Mem_Nov2020/MSMemNov2020_R/MSMem_Github/R_MSMem.csv")

# subsets 
po_data<-subset(data, Surgery == "LTLAmHipp" | Surgery == "RTLAmHipp" ) # excluding controls and selective surgery patients 

```

# Group differences in degree left ATL

Z-score across groups at each threshold, average the z-scored degree across thresholds, and then correlate 

## Prepare the data 
```{r}

# z-score across the groups at each threshold 
for (cost in c("05","10","15","20","25","30","35","40","45")){
  eval(parse(text = paste0("data$L_cost",cost,"_z<-scale(data$L_cost",cost,")")))
  }

# mean acros thresholds 
data$mean_L_degree<-rowMeans(data[ , c("L_cost05_z","L_cost10_z","L_cost15_z","L_cost20_z","L_cost25_z","L_cost30_z","L_cost35_z","L_cost40_z","L_cost45_z")])

```

## Models 
```{r}
# model 
model<-lm(mean_L_degree ~ Group, data=data)

## test model assumptions 
plot(model) # residuals vs fitted should be straight, no bad outliers 
hist(residuals(model)) # residuals should be normal 
shapiro.test(residuals(model)) # residuals should be normal
ncvTest(model) # test for heterogeneity of variance, it's violated here 

# since the model was heteroskedastic, run welch anova instead 
welch_anova_test(mean_L_degree ~ Group, data=data) 

# post hoc test 
games_howell_test(mean_L_degree ~ Group, data=data) # using games howell since homogeneity of variance is violated
```

## Plot 
```{r, fig.height = 2, fig.width=2}
# plot 
ggplot(data, aes(x=Group, y=mean_L_degree, fill=Group)) +
  geom_boxplot() +
  geom_point(position=position_jitterdodge(jitter.width=0.5, dodge.width = 0),
             pch=21, aes(fill=factor(Group)), show.legend = F, size = 1.5) +
  scale_y_continuous(breaks = seq(-2,3,1),expand = c(0, 0), limits = c(-2,3))+
  theme_classic() +
  theme(text = element_text(size = 23), plot.title = element_text(hjust = 0.5), legend.position = "none") +
  scale_fill_brewer(palette="Blues") +
  labs(x = "",
     y = "mean left ATL degree (z)",
     title = "")

```



# Group differences in degree right ATL

Z-score across groups at each threshold, average the z-scored degree across thresholds, and then correlate 

## Prepare the data
```{r}

# z-score across the groups at each threshold 
for (cost in c("05","10","15","20","25","30","35","40","45")){
  eval(parse(text = paste0("data$R_cost",cost,"_z<-scale(data$R_cost",cost,")")))
}

# average across thresholds 
data$mean_R_degree<-rowMeans(data[ , c("R_cost05_z","R_cost10_z","R_cost15_z","R_cost20_z","R_cost25_z","R_cost30_z","R_cost35_z","R_cost40_z","R_cost45_z")])

```

## Models 
```{r}
# model 
model<-lm(mean_R_degree ~ Group, data=data)

# check model assumptions 
plot(model) # residuals vs fitted should be straight, no bad outliers 
hist(residuals(model)) # residuals should be normal 
shapiro.test(residuals(model)) # residuals should be normal
ncvTest(model) # test for heterogeneity of variance

# anova
Anova(model, type="III")

# post hocs 
pairwise<-emmeans(model, list(pairwise ~ Group), adjust = "none")
pairwise
confint(pairwise, level=0.95, oldNames=FALSE)

```

## Plot
```{r, fig.height = 2, fig.width=2}
# plot 
ggplot(data, aes(x=Group, y=mean_R_degree, fill=Group)) +
  geom_boxplot() +
  geom_point(position=position_jitterdodge(jitter.width=0.5, dodge.width = 0),
             pch=21, aes(fill=factor(Group)), show.legend = F, size = 1.5) +
  scale_y_continuous(breaks = seq(-2,3,1),expand = c(0, 0), limits = c(-2,3))+
  theme_classic() +
  theme(text = element_text(size = 23), plot.title = element_text(hjust = 0.5), legend.position = "none") +
  scale_fill_brewer(palette="Blues") +
  labs(x = "",
     y = "mean right ATL degree (z)",
     title = "")

```


# Test relationship between degree and postop memory outcomes 

Excluding selective surgery patients 

## Prepare the data 
```{r}

# z score across thresholds, for each hemisphere 
for (cost in c("05","10","15","20","25","30","35","40","45")){
  for (hem in c("L","R")){
    eval(parse(text = paste0("po_data$",hem,"_cost",cost,"_z<-scale(po_data$",hem,"_cost",cost,")")))
  }
}

# average across thresholds 
po_data$mean_L_degree<-rowMeans(po_data[ , c("L_cost05_z","L_cost10_z","L_cost15_z","L_cost20_z","L_cost25_z","L_cost30_z","L_cost35_z","L_cost40_z","L_cost45_z")])

po_data$mean_R_degree<-rowMeans(po_data[ , c("R_cost05_z","R_cost10_z","R_cost15_z","R_cost20_z","R_cost25_z","R_cost30_z","R_cost35_z","R_cost40_z","R_cost45_z")])

# create a variable that reflects degree in left ATL for LTLE and right ATL for RTLE
po_data<-po_data%>%
  mutate(mean_ATL_degree = case_when(Group == "LTLE" ~ mean_L_degree,
            Group == "RTLE" ~ mean_R_degree))
```

## Model
```{r}
# model 
model<-lm(material_specific_mem_change ~ mean_ATL_degree*Group, data=po_data)

# test model assumptions 
plot(model) # residuals vs fitted should be straight, no bad outliers 
hist(residuals(model)) # residuals should be normal 
shapiro.test(residuals(model)) # residuals should be normal
ncvTest(model) # test for heterogeneity of variance 
vif(model) # collinearity check, should be less than 5 

# regression 
summary(model)
confint(model)

## post hoc test of simple slopes 
emtrends(model, ~Group, var="mean_ATL_degree")
## trend is the slope. Can see based on confidence intervals RTLE is crossing zero=ns, but LTLE is sig 

```

# Plots

## Degree relationship with material specific memory, averaged across thresholds
```{r, fig.height = 4, fig.width=4}

# subset the data 
LTLE_po_data<-subset(po_data, Group=="LTLE")
RTLE_po_data<-subset(po_data, Group=="RTLE")


# plot the summary correlation 
LTLE_po_data %>%
  ggplot(aes(x = mean_L_degree, y = VerbalFactor)) +
  geom_smooth(method = lm, size = 2, alpha = 0.2, level = 0.95, colour="black") +
  geom_point(size = 3.5) +
  theme_classic() +
  theme(text = element_text(size = 23), plot.title = element_text(hjust=0.5, face = "bold", size = 22))+
  scale_y_continuous(limits = c(-3,3), breaks = seq(-3,3,1))+
  scale_x_continuous(limits = c(-3,2), breaks = seq(-3,2,1))+
  labs(x = "mean LTBR degree (z)",
       y = "verbal memory change",
       title = "")

# plot the summary correlation 
RTLE_po_data %>%
  ggplot(aes(x = mean_R_degree, y = VisualFactor)) +
  geom_smooth(method = lm, size = 2, alpha = 0.2, level = 0.95, colour="black") +
  geom_point(size = 3.5) +
  theme_classic() +
  theme(text = element_text(size = 23), plot.title = element_text(hjust=0.5, face = "bold", size = 22))+
  scale_y_continuous(limits = c(-3,3), breaks = seq(-3,3,1))+
  scale_x_continuous(limits = c(-3,2), breaks = seq(-3,2,1))+
  labs(x = "mean RTBR degree (z)",
       y = "vis. memory change",
       title = "")




```
## again with hemisphere dominance coloured in 
To check that language dominance isn't driving our correlations. 
```{r, fig.height = 5, fig.width=6}

RTLE_po_data %>%
  ggplot(aes(x = mean_R_degree, y = VisualFactor, color = hem_dom)) +
  geom_smooth(method = lm, size = 2, alpha = 0.2, level = 0.95, colour="black") +
  geom_point(size = 3.5) +
  theme_classic() +
  theme(text = element_text(size = 23), plot.title = element_text(hjust=0.5, face = "bold", size = 22))+
  scale_y_continuous(limits = c(-3,3), breaks = seq(-3,3,1))+
  scale_x_continuous(limits = c(-3,2), breaks = seq(-3,2,1))+
  scale_color_manual(values=c("#3ebbfa","#fc5e03")) +
  labs(x = "mean RTBR degree (z)",
       y = "vis. memory change",
       title = "")

LTLE_po_data %>%
  ggplot(aes(x = mean_L_degree, y = VerbalFactor, color = hem_dom)) +
  geom_smooth(method = lm, size = 2, alpha = 0.2, level = 0.95, colour="black") +
  geom_point(size = 3.5) +
  theme_classic() +
  theme(text = element_text(size = 23), plot.title = element_text(hjust=0.5, face = "bold", size = 22))+
  scale_y_continuous(limits = c(-3,3), breaks = seq(-3,3,1))+
  scale_x_continuous(limits = c(-3,2), breaks = seq(-3,2,1))+
  scale_color_manual(values=c("#3ebbfa","#fc5e03")) +
  labs(x = "mean LTBR degree (z)",
       y = "verbal memory change",
       title = "")

```



## LTLE correlations between degree and verbal memory change across thresholds 
Check that no one threshold is driving the average correlation. 
```{r, fig.height = 4, fig.width=4}

##correlate degree across different thresholds with memory change. Using one tailed tests given our prior work and hypotheses
attach(LTLE_po_data)
L_cost05<-cor.test(VerbalFactor, L_cost05, alternative="less", data=LTLE_po_data)
L_cost10<-cor.test(VerbalFactor, L_cost10, alternative="less", data=LTLE_po_data)
L_cost15<-cor.test(VerbalFactor, L_cost15, alternative="less", data=LTLE_po_data)
L_cost20<-cor.test(VerbalFactor, L_cost20, alternative="less", data=LTLE_po_data)
L_cost25<-cor.test(VerbalFactor, L_cost25, alternative="less", data=LTLE_po_data)
L_cost30<-cor.test(VerbalFactor, L_cost30, alternative="less", data=LTLE_po_data)
L_cost35<-cor.test(VerbalFactor, L_cost35, alternative="less", data=LTLE_po_data)
L_cost40<-cor.test(VerbalFactor, L_cost40, alternative="less", data=LTLE_po_data)
L_cost45<-cor.test(VerbalFactor, L_cost45, alternative="less", data=LTLE_po_data)

## create dataframe 
L_cost_correlation_CI <- data.frame(
  upper= c(L_cost05$conf.int[2],L_cost10$conf.int[2], L_cost15$conf.int[2], L_cost20$conf.int[2], L_cost25$conf.int[2], L_cost30$conf.int[2], L_cost35$conf.int[2],L_cost40$conf.int[2], L_cost45$conf.int[2]),
  lower = c(L_cost05$conf.int[1],L_cost10$conf.int[1], L_cost15$conf.int[1], L_cost20$conf.int[1], L_cost25$conf.int[1], L_cost30$conf.int[1], L_cost35$conf.int[1],L_cost40$conf.int[1], L_cost45$conf.int[1] ),
  correlation =c(L_cost05$estimate[1],L_cost10$estimate[1], L_cost15$estimate[1], L_cost20$estimate[1], L_cost25$estimate[1], L_cost30$estimate[1], L_cost35$estimate[1],L_cost40$estimate[1], L_cost45$estimate[1]),
  threshold = c(5, 10, 15,20,25,30,35,40,45)
)

# plot the correlations across thresholds 
L_cost_correlation_CI %>%
  ggplot(aes( x = threshold, y= correlation)) +
  geom_line(color = "black", size = 1)+
  geom_pointrange(aes(ymin = lower, ymax = upper),color = "black", size = 1)+
  geom_abline(aes(intercept = 0, slope = 0), linetype = 2) +
  theme_classic()+
  theme(text = element_text(size = 22), plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(limits = c(-1, 0.5), breaks=seq(-1,0.5,0.2)) +
  scale_x_continuous(limits = c(5, 45), breaks=seq(5,45,5)) +
  labs(y = "LATL degree corr. with
       verbal memory change",
       x = "threshold",
       title = "")

detach(LTLE_po_data)


```

## RTLE correlations between degree and visual memory change across thresholds 
```{r, fig.height = 4, fig.width=4}

##correlate degree across different thresholds with memory change. Using one tailed tests given our prior work and hypotheses
attach(RTLE_po_data)
R_cost05<-cor.test(VisualFactor, R_cost05, alternative="less", data=RTLE_po_data)
R_cost10<-cor.test(VisualFactor, R_cost10, alternative="less", data=RTLE_po_data)
R_cost15<-cor.test(VisualFactor, R_cost15, alternative="less", data=RTLE_po_data)
R_cost20<-cor.test(VisualFactor, R_cost20, alternative="less", data=RTLE_po_data)
R_cost25<-cor.test(VisualFactor, R_cost25, alternative="less", data=RTLE_po_data)
R_cost30<-cor.test(VisualFactor, R_cost30, alternative="less", data=RTLE_po_data)
R_cost35<-cor.test(VisualFactor, R_cost35, alternative="less", data=RTLE_po_data)
R_cost40<-cor.test(VisualFactor, R_cost40, alternative="less", data=RTLE_po_data)
R_cost45<-cor.test(VisualFactor, R_cost45, alternative="less", data=RTLE_po_data)

## create data frame 
R_cost_correlation_CI <- data.frame(
  upper= c(R_cost05$conf.int[2],R_cost10$conf.int[2], R_cost15$conf.int[2], R_cost20$conf.int[2], R_cost25$conf.int[2], R_cost30$conf.int[2], R_cost35$conf.int[2],R_cost40$conf.int[2], R_cost45$conf.int[2]),
  lower = c(R_cost05$conf.int[1],R_cost10$conf.int[1], R_cost15$conf.int[1], R_cost20$conf.int[1], R_cost25$conf.int[1], R_cost30$conf.int[1], R_cost35$conf.int[1],R_cost40$conf.int[1], R_cost45$conf.int[1] ),
  correlation =c(R_cost05$estimate[1],R_cost10$estimate[1], R_cost15$estimate[1], R_cost20$estimate[1], R_cost25$estimate[1], R_cost30$estimate[1], R_cost35$estimate[1],R_cost40$estimate[1], R_cost45$estimate[1]),
  threshold = c(5, 10, 15,20,25,30,35,40,45)
)

## plot the correlation 
R_cost_correlation_CI %>%
  ggplot(aes( x = threshold, y= correlation)) +
  geom_line(color = "black", size = 1)+
  geom_pointrange(aes(ymin = lower, ymax = upper),color = "black", size = 1)+
  geom_abline(aes(intercept = 0, slope = 0), linetype = 2) +
  theme_classic()+
  theme(text = element_text(size = 23), plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(limits = c(-1, 0.5), breaks=seq(-1,0.5,0.2)) +
  scale_x_continuous(limits = c(5, 45), breaks=seq(5,45,5)) +
  labs(y = "RATL degree corr. with 
       vis. memory change",
       x = "threshold",
       title = "")

```



Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

