---
title: "MSMem Matrix Similarity and ROI Similarity Analyses"
output: html_notebook
---

V1.0 

Sam Audrain, December 2022

This script takes already calculated matrix similarity values and uses them to predict material-specific memory change in left and right TLE groups. 

It also correlates already computed ROI-similarity values with material-specific memory change in each group. 


# Read in the packages

```{r}

library(dplyr)
library(ggplot2)
library(emmeans)
library(rstatix)
library(car)


```

# Read in the data 
```{r}
###### CHANGE PATH FOR YOUR DATA #######
data<-read.csv("/Users/audrainsp/Library/Mobile Documents/com~apple~CloudDocs/Documents/Science_Projects/RSMemory/MatrixSim_Mem_Nov2020/MSMemNov2020_R/MSMem_Github/R_MSMem.csv")

# subset the data
po_data<-subset(data, material_specific_mem_change !="NA")
LTLE_po_data<-subset(data, Group=="LTLE")
RTLE_po_data<-subset(data, Group=="RTLE")

```


# Matrix Similarity Regression 
```{r, fig.height=5,fig.width=5}

# model 
model<-lm(material_specific_mem_change ~ MatrixSim_64 * Group, data=po_data, na.action=na.omit)

# test model assumptions 
plot(model) # residuals vs fitted should be straight, no bad outliers 
hist(residuals(model)) # residuals should be normal 
shapiro.test(residuals(model)) # residuals should be normal 
ncvTest(model) # test for heterogeneity of variance 

# anova table 
summary(model)

## post hoc test of simple slopes 
emtrends(model, ~Group, var="MatrixSim_64")
## trend is the slope. Can see based on confidence intervals RTLE is crossing zero=ns, but LTLE is sig 
```

# Plots 
```{r, fig.height=5,fig.width=5}

#### LTLE ######
LTLE_po_data %>%
  ggplot(aes(y = VerbalFactor, x = MatrixSim_64)) +
  geom_smooth(method = lm, size = 2, alpha = 0.2, level = 0.95, colour="black") +
  geom_point(size = 3.5) +
  theme_classic() +
  theme(text = element_text(size = 25), plot.title = element_text(hjust=0.5, face = "bold", size = 22))+
  scale_y_continuous(limits = c(-3,3), breaks = seq(-3,3,1))+
  labs(y = "verbal memory change",
       x = "matrix similarity",
       title = "")

## RTLE 
RTLE_po_data %>%
  ggplot(aes(y = VisualFactor, x = MatrixSim_64)) +
  geom_smooth(method = lm, size = 2, alpha = 0.2, level = 0.95, colour="black") +
  geom_point(size = 3.5) +
  theme_classic() +
  theme(text = element_text(size = 25), plot.title = element_text(hjust=0.5, face = "bold", size = 22))+
  scale_y_continuous(limits = c(-3,3), breaks = seq(-3,3,1))+
  labs(y = "visuospatial memory change",
       x = "matrix similarity",
       title = "")
 
```

# ROI-similarity Correlations 

Mass correlations for ROI similarity and p-value correction 

## LTLE

This correlates ROI-similarity with verbal memory change in LTLE, and corrects for multiple comparisons. 

Resulting variable pvals_adjusted_LTLE contains adjusted p values for each BNA_ROI in the network (each row is a p value for 1 ROI, output in the same order as they appear in the R_MsMem.csv spreadsheet)
```{r}

pvals<-0

## 18:81 indexes columns that contain ROI-similarity values for each BNA_ROI, and each participant (the rows)
for (i in 18:81) {
  a <- cor.test(LTLE_po_data$VerbalFactor, LTLE_po_data[,i], method = "pearson", alternative = "two.sided")
  #print(a$estimate)
  #print(a$p.value)
  pvals<-rbind(pvals,a$p.value)
}

# FDR correct pvalues 
pvals<-as.matrix(pvals[-1,])
pvals_adjusted_LTLE<-as.matrix(p.adjust(pvals, method = "fdr", n = length(pvals))) 

```

## RTLE

This correlates ROI-similarity with visual memory change in RTLE, and corrects for multiple comparisons. 

Resulting variable pvals_adjusted_RTLE contains adjusted p values for each BNA_ROI in the network (each row is a p value for 1 ROI, output in the same order as they appear in the R_MsMem.csv spreadsheet)

```{r}

### Rights  ###
pvals<-0
## 18:81 indexes columns that contain ROI-similarity values for each BNA_ROI, and each participant (the rows)
for (i in 18:81) {
  a <- cor.test(RTLE_po_data$VisualFactor, RTLE_po_data[,i], method = "pearson", alternative = "two.sided")
  #print(a$estimate)
  #print(a$p.value)
  pvals<-rbind(pvals,a$p.value)
}

# FDR correct pvalues 
pvals<-pvals[-1,]
pvals_adjusted_RTLE<-as.matrix(p.adjust(pvals, method = "fdr", n = length(pvals))) 

```



Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

