---
title: "MSMem Multiple Regression Model Comparisons"
output: html_notebook
---

V2.0 edits based on peer review 

Sam Audrain, September 2023

Analyzes variance explained by Matrix Similarity and Mean ATL Degree predictors and compares to clinical and demographic variables. 
 
# Read in the packages
```{r}

library(car)
library(jtools)
library(dplyr)
library(ggplot2)
library(emmeans)

```

# Read in the data
```{r}
###### CHANGE PATH FOR YOUR DATA #######
data<-read.csv("/Users/samanthaaudrain/Documents/Science_Projects/RSMemory/MatrixSim_Mem_Nov2020/MSMemNov2020_R/MsMem_Github/R_MSMem.csv")

# subset 
po_data<-subset(data, Surgery == "LTLAmHipp" | Surgery == "RTLAmHipp" ) # excluding controls and selective surgery patients 

```


# Prepare the data 

To statistically test, we will z-score across groups at each threshold, average the z-scored degree across thresholds, and then correlate once

```{r}

# z score across thresholds, for each hemisphere 
for (cost in c("05","10","15","20","25","30","35","40","45")){
  for (hem in c("L","R")){
    eval(parse(text = paste0("po_data$",hem,"_cost",cost,"_z<-scale(po_data$",hem,"_cost",cost,")")))
  }
}

# average across thresholds 
po_data$mean_L_degree<-rowMeans(po_data[ , c("L_cost05_z","L_cost10_z","L_cost15_z","L_cost20_z","L_cost25_z","L_cost30_z","L_cost35_z","L_cost40_z","L_cost45_z")])

po_data$mean_R_degree<-rowMeans(po_data[ , c("R_cost05_z","R_cost10_z","R_cost15_z","R_cost20_z","R_cost25_z","R_cost30_z","R_cost35_z","R_cost40_z","R_cost45_z")])

## center MS
po_data$MatrixSim_64_scaled<-scale(po_data$MatrixSim_64, center=TRUE, scale=FALSE)

cor.test(po_data$mean_L_degree, po_data$MatrixSim_64)

```

# LTLE

## Models

### Verbal memory change predicted by matrix similarity + degree
```{r}

# model 
model<-lm(VerbalFactor ~ MatrixSim_64_scaled + mean_L_degree, data=po_data, na.action=na.omit, subset=Group=="LTLE")

# check model assumptions 
plot(model) # residuals vs fitted should be straight, no bad outliers
hist(residuals(model)) # residuals should be normal
shapiro.test(residuals(model)) # residuals should be normal
ncvTest(model) # test for heterogeneity of variance
vif(model) ## variance inflation should be less than 5-10 

# stats 
summary(model) ##when using just full ATL cases, both degree and MS predicts outcome
confint(model, level=0.95)

# variance explained 
summ(model, part.corr=TRUE) ### part.r is semipartial corr, partial.r is partial corr. 
#semi-partial correlation = does A influence B above and beyond C influence on B? 
#partial correlation = what is the correlation between A and B after accounting for fact that C influences both? 
## to get unique variance, we want semi-partial corrs. to get shared variance, subtract sub of semi-partial from total model r2 
# remember these are rs, need to square to get % variance explained 
```

#### Testing effect of days since surgery on MS and ROIsim 
```{r}

# model 
model3<-lm(VerbalFactor ~ MatrixSim_64_scaled + mean_L_degree + surg_NP2_delay + MatrixSim_64_scaled*surg_NP2_delay + mean_L_degree*surg_NP2_delay, data=po_data, na.action=na.omit, subset=Group=="LTLE")

# check model assumptions 
plot(model3) # residuals vs fitted should be straight, no bad outliers
hist(residuals(model3)) # residuals should be normal
shapiro.test(residuals(model3)) # residuals should be normal
ncvTest(model3) # test for heterogeneity of variance
vif(model3, type='predictor') ## variance inflation should be less than 5-10 

# stats 
summary(model3) ##when using just full ATL cases, both degree and MS predicts outcome
confint(model3, level=0.95)

# posthoc tests of interactions 
### subset the data 
po_data_LTLE<-subset(po_data, Group=="LTLE") 

### get high medium and low values of our continuous variable, surgery delay 
delay_high<-mean(po_data_LTLE$surg_NP2_delay)+sd(po_data_LTLE$surg_NP2_delay)
delay_med<-mean(po_data_LTLE$surg_NP2_delay)
delay_low<-mean(po_data_LTLE$surg_NP2_delay)-sd(po_data_LTLE$surg_NP2_delay)

### emtrends to test 
emtrends(model3, pairwise ~ surg_NP2_delay, var="mean_L_degree", at=list(surg_NP2_delay=c(delay_high,delay_med,delay_low)), adjust="none", data=po_data_LTLE)


```


### Verbal memory change predicted by matrix similarity + degree + clinical and demographic variables 
```{r}
## model 
model2<-lm(VerbalFactor ~ MatrixSim_64_scaled + mean_L_degree + Pre_Verb + LH_vol_norm +age1 +sex +edu +AOO, data=po_data, na.action=na.omit, subset=Group=="LTLE")

# check model assumptions 
plot(model2) # residuals vs fitted should be straight, no bad outliers
hist(residuals(model2)) # residuals should be normal
shapiro.test(residuals(model2)) # residuals should be normal
ncvTest(model2) # test for heterogeneity of variance
vif(model2) ## variance inflation should be less than 5-10 

# stats 
summary(model2) ##when using just full ATL cases, both degree and MS predicts outcome 
confint(model2, level=0.95)

# variance explained 
summ(model2, part.corr=TRUE) ### part.r is semipartial corr, partial.r is partial corr. 
#semi-partial correlation = does A influence B above and beyond C influence on B? 
#partial correlation = what is the correlation between A and B after accounting for fact that C influences both? 
## to get unique variance, we want semi-partial corrs. to get shared variance, subtract sub of semi-partial from total model r2 
# remember these are rs, need to square to get % variance explained 

```

### Compare models with and without clinical and demographic variables 
```{r}

anova(model,model2)

```

### Exploratory correlations between clinical variables and verbal postop change 

```{r, fig.height=5,fig.width=5.5}

po_data_LTLE<-subset(po_data, Group=="LTLE")
po_data_RTLE<-subset(po_data, Group=="RTLE")

po_data %>%
  ggplot(aes(y = VerbalFactor, x = Pre_Verb, group = Group, color = Group)) +
  geom_smooth(method = lm, size = 2, alpha = 0.2, level = 0.95) +
  geom_point(size = 3.5) +
  theme_classic() +
  theme(text = element_text(size = 25), plot.title = element_text(hjust=0.5, face = "bold", size = 22))+
  scale_y_continuous(limits = c(-3,3), breaks = seq(-3,3,1))+
  scale_color_manual(values=c("#3ebbfa","#fc5e03")) +
  labs(y = "verbal memory change",
       x = "preop verbal memory",
       title = "")

cor.test(po_data_LTLE$Pre_Verb, po_data_LTLE$VerbalFactor)
cor.test(po_data_RTLE$Pre_Verb, po_data_RTLE$VerbalFactor)

po_data %>%
  ggplot(aes(y = VerbalFactor, x = age1, group = Group, color = Group)) +
  geom_smooth(method = lm, size = 2, alpha = 0.2, level = 0.95) +
  geom_point(size = 3.5) +
  theme_classic() +
  theme(text = element_text(size = 25), plot.title = element_text(hjust=0.5, face = "bold", size = 22))+
  scale_y_continuous(limits = c(-3,3), breaks = seq(-3,3,1))+
  scale_color_manual(values=c("#3ebbfa","#fc5e03")) +
  labs(y = "verbal memory change",
       x = "age",
       title = "")

cor.test(po_data_LTLE$age1, po_data_LTLE$VerbalFactor)
cor.test(po_data_RTLE$age1, po_data_RTLE$VerbalFactor)

po_data %>%
  ggplot(aes(y = VerbalFactor, x = edu, group = Group, color = Group)) +
  geom_smooth(method = lm, size = 2, alpha = 0.2, level = 0.95) +
  geom_point(size = 3.5) +
  theme_classic() +
  theme(text = element_text(size = 25), plot.title = element_text(hjust=0.5, face = "bold", size = 22))+
  scale_y_continuous(limits = c(-3,3), breaks = seq(-3,3,1))+
  scale_color_manual(values=c("#3ebbfa","#fc5e03")) +
  labs(y = "verbal memory change",
       x = "education",
       title = "")

cor.test(po_data_LTLE$edu, po_data_LTLE$VerbalFactor)
cor.test(po_data_RTLE$edu, po_data_RTLE$VerbalFactor)


po_data %>%
  ggplot(aes(y = VerbalFactor, x = AOO, group = Group, color = Group)) +
  geom_smooth(method = lm, size = 2, alpha = 0.2, level = 0.95) +
  geom_point(size = 3.5) +
  theme_classic() +
  theme(text = element_text(size = 25), plot.title = element_text(hjust=0.5, face = "bold", size = 22))+
  scale_y_continuous(limits = c(-3,3), breaks = seq(-3,3,1))+
  scale_color_manual(values=c("#3ebbfa","#fc5e03")) +
  labs(y = "verbal memory change",
       x = "age of onset",
       title = "")

cor.test(po_data_LTLE$AOO, po_data_LTLE$VerbalFactor)
cor.test(po_data_RTLE$AOO, po_data_RTLE$VerbalFactor)


po_data %>%
  ggplot(aes(x=sex, y=VerbalFactor, fill=Group)) +
    geom_boxplot() +
    geom_point(position=position_jitterdodge(jitter.width=0.3, dodge.width = 0.8),
               pch=21, aes(fill=factor(Group)), show.legend = F, size = 1.5) +
    scale_y_continuous(breaks = seq(-3,3,1),expand = c(0, 0), limits = c(-3,3))+
    theme_classic() +
    theme(text = element_text(size = 23), plot.title = element_text(hjust = 0.5), legend.position = "right") +
    scale_fill_manual(values=c("#3ebbfa","#fc5e03")) +
    labs(y = "verbal memory change",
         x = "sex",
         title = "")

t.test(VerbalFactor ~ sex, data=po_data_LTLE)
t.test(VerbalFactor ~ sex, data=po_data_RTLE)


### WILL BE REPALCED BY HIPPO VOLUME 
po_data %>%
  ggplot(aes(x=as.factor(MTS_status), y=VerbalFactor, fill=Group)) +
    geom_boxplot() +
    geom_point(position=position_jitterdodge(jitter.width=0.3, dodge.width = 0.8),
               pch=21, aes(fill=factor(Group)), show.legend = F, size = 1.5) +
    scale_y_continuous(breaks = seq(-3,3,1),expand = c(0, 0), limits = c(-3,3))+
    theme_classic() +
    theme(text = element_text(size = 23), plot.title = element_text(hjust = 0.5), legend.position = "right") +
    scale_fill_manual(values=c("#3ebbfa","#fc5e03")) +
    scale_x_discrete(labels=c("0" = "no MTS", "1" = "MTS")) +
    labs(y = "verbal memory change",
         x = "MTS",
         title = "")

t.test(VerbalFactor ~ MTS_status, data=po_data_LTLE)
t.test(VerbalFactor ~ MTS_status, data=po_data_RTLE)


po_data %>%
  ggplot(aes(y = VerbalFactor, x = LH_vol_norm, group = Group, color = Group)) +
  geom_smooth(method = lm, size = 2, alpha = 0.2, level = 0.95) +
  geom_point(size = 3.5) +
  theme_classic() +
  theme(text = element_text(size = 25), plot.title = element_text(hjust=0.5, face = "bold", size = 22))+
  scale_y_continuous(limits = c(-3,3), breaks = seq(-3,3,1))+
  scale_color_manual(values=c("#3ebbfa","#fc5e03")) +
  labs(y = "verbal memory change",
       x = "left hippocampal volume",
       title = "")

cor.test(po_data_LTLE$LH_vol_norm, po_data_LTLE$VerbalFactor)
cor.test(po_data_RTLE$LH_vol_norm, po_data_RTLE$VerbalFactor)


```





# RTLE

## Models

### Visual memory change predicted by clinical and demographic variables

We already know matrix similarity and degree are not associated with visual memory change, so we don't include them here. 

```{r}
# model 
model<-lm(VisualFactor ~ Pre_Vis +RH_vol_norm +age1 +sex +edu +AOO, data=po_data, na.action=na.omit, subset=Group=="RTLE") 

# check model assumptions 
plot(model2) # residuals vs fitted should be straight, no bad outliers
hist(residuals(model2)) # residuals should be normal
shapiro.test(residuals(model2)) # residuals should be normal
ncvTest(model2) # test for heterogeneity of variance
vif(model2) ## variance inflation should be less than 10 

# stats 
summary(model)


```

### Correlations between clinical variables and visual postop change 

```{r, fig.height=5,fig.width=5.5}

po_data %>%
  ggplot(aes(y = VisualFactor, x = Pre_Vis, group = Group, color = Group)) +
  geom_smooth(method = lm, size = 2, alpha = 0.2, level = 0.95) +
  geom_point(size = 3.5) +
  theme_classic() +
  theme(text = element_text(size = 25), plot.title = element_text(hjust=0.5, face = "bold", size = 22))+
  scale_y_continuous(limits = c(-3,3), breaks = seq(-3,3,1))+
  scale_color_manual(values=c("#3ebbfa","#fc5e03")) +
  labs(y = "visuospatial memory change",
       x = "preop visuospatial memory",
       title = "")

cor.test(po_data_LTLE$Pre_Vis, po_data_LTLE$VisualFactor)
cor.test(po_data_RTLE$Pre_Vis, po_data_RTLE$VisualFactor)

po_data %>%
  ggplot(aes(y = VisualFactor, x = age1, group = Group, color = Group)) +
  geom_smooth(method = lm, size = 2, alpha = 0.2, level = 0.95) +
  geom_point(size = 3.5) +
  theme_classic() +
  theme(text = element_text(size = 25), plot.title = element_text(hjust=0.5, face = "bold", size = 22))+
  scale_y_continuous(limits = c(-3,3), breaks = seq(-3,3,1))+
  scale_color_manual(values=c("#3ebbfa","#fc5e03")) +
  labs(y = "visuospatial memory change",
       x = "age",
       title = "")

cor.test(po_data_LTLE$age1, po_data_LTLE$VisualFactor)
cor.test(po_data_RTLE$age1, po_data_RTLE$VisualFactor)

po_data %>%
  ggplot(aes(y = VisualFactor, x = edu, group = Group, color = Group)) +
  geom_smooth(method = lm, size = 2, alpha = 0.2, level = 0.95) +
  geom_point(size = 3.5) +
  theme_classic() +
  theme(text = element_text(size = 25), plot.title = element_text(hjust=0.5, face = "bold", size = 22))+
  scale_y_continuous(limits = c(-3,3), breaks = seq(-3,3,1))+
  scale_color_manual(values=c("#3ebbfa","#fc5e03")) +
  labs(y = "visuospatial memory change",
       x = "education",
       title = "")

cor.test(po_data_LTLE$edu, po_data_LTLE$VisualFactor)
cor.test(po_data_RTLE$edu, po_data_RTLE$VisualFactor)

po_data %>%
  ggplot(aes(y = VisualFactor, x = AOO, group = Group, color = Group)) +
  geom_smooth(method = lm, size = 2, alpha = 0.2, level = 0.95) +
  geom_point(size = 3.5) +
  theme_classic() +
  theme(text = element_text(size = 25), plot.title = element_text(hjust=0.5, face = "bold", size = 22))+
  scale_y_continuous(limits = c(-3,3), breaks = seq(-3,3,1))+
  scale_color_manual(values=c("#3ebbfa","#fc5e03")) +
  labs(y = "visuospatial memory change",
       x = "age of onset",
       title = "")

cor.test(po_data_LTLE$AOO, po_data_LTLE$VisualFactor)
cor.test(po_data_RTLE$AOO, po_data_RTLE$VisualFactor)

po_data %>%
  ggplot(aes(x=sex, y=VisualFactor, fill=Group)) +
    geom_boxplot() +
    geom_point(position=position_jitterdodge(jitter.width=0.3, dodge.width = 0.8),
               pch=21, aes(fill=factor(Group)), show.legend = F, size = 1.5) +
    scale_y_continuous(breaks = seq(-3,3,1),expand = c(0, 0), limits = c(-3,3))+
    theme_classic() +
    theme(text = element_text(size = 23), plot.title = element_text(hjust = 0.5), legend.position = "right") +
    #scale_fill_brewer(palette="Blues") +
    scale_fill_manual(values=c("#3ebbfa","#fc5e03")) +
    labs(y = "visuospatial memory change",
         x = "sex",
         title = "")

t.test(VisualFactor ~ sex, data=po_data_LTLE)
t.test(VisualFactor ~ sex, data=po_data_RTLE)

po_data %>%
  ggplot(aes(x=as.factor(MTS_status), y=VisualFactor, fill=Group)) +
    geom_boxplot() +
    geom_point(position=position_jitterdodge(jitter.width=0.3, dodge.width = 0.8),
               pch=21, aes(fill=factor(Group)), show.legend = F, size = 1.5) +
    scale_y_continuous(breaks = seq(-3,3,1),expand = c(0, 0), limits = c(-3,3))+
    theme_classic() +
    theme(text = element_text(size = 23), plot.title = element_text(hjust = 0.5), legend.position = "right") +
    scale_fill_manual(values=c("#3ebbfa","#fc5e03")) +
    scale_x_discrete(labels=c("0" = "no MTS", "1" = "MTS")) +
    labs(y = "visuospatial memory change",
         x = "MTS",
         title = "")

po_data %>%
  ggplot(aes(y = VisualFactor, x = RH_vol_norm, group = Group, color = Group)) +
  geom_smooth(method = lm, size = 2, alpha = 0.2, level = 0.95) +
  geom_point(size = 3.5) +
  theme_classic() +
  theme(text = element_text(size = 25), plot.title = element_text(hjust=0.5, face = "bold", size = 22))+
  scale_y_continuous(limits = c(-3,3), breaks = seq(-3,3,1))+
  scale_color_manual(values=c("#3ebbfa","#fc5e03")) +
  labs(y = "visuospatial memory change",
       x = "right hippocampal volume",
       title = "")

cor.test(po_data_LTLE$RH_vol_norm, po_data_LTLE$VisualFactor)
cor.test(po_data_RTLE$RH_vol_norm, po_data_RTLE$VisualFactor)

```

