---
title: "MSMem Demographic Analysis"
output: html_notebook
---
V1.0 

Sam Audrain, December 2022

This script tests for statistical differences in demographic and clinical variables between right TLE, left TLE, and control groups. 
 
# Read in packages
```{r}

library(dplyr)
library(car)
library(rstatix)

```

# Read and subset the data 
```{r}
###### CHANGE PATH FOR YOUR DATA #######
data_demos<-read.csv("/Users/audrainsp/Library/Mobile Documents/com~apple~CloudDocs/Documents/Science_Projects/RSMemory/MatrixSim_Mem_Nov2020/MSMemNov2020_R/MSMem_Github/R_MSMem.csv")

# subset
demos_TLE<-subset(data_demos, Group == "LTLE" | Group == "RTLE")
demos_LTLE<-subset(data_demos, Group == "LTLE")
demos_RTLE<-subset(data_demos, Group == "RTLE")

```

# Age 
```{r}
# summary per group 
data_demos %>%
  group_by(Group) %>%
  summarize(mean = mean(age1, na.rm = TRUE), sd = sd(age1, na.rm = TRUE), min = min(age1, na.rm = TRUE), max = max(age1, na.rm = TRUE), n = n())

# model 
model<-lm(age1~Group, data=data_demos, na.action=na.omit) 

# check model assumptions 
plot(model) # residuals vs fitted should be straight, no bad outliers 
hist(residuals(model)) # residuals should be normal 
shapiro.test(residuals(model)) # residuals should be normal 
ncvTest(model) # test for heterogeneity of variance 

# normality assumption violated 
data_demos %>%
  group_by(Group) %>%
  shapiro_test(age1)

# nonparametric test 
kruskal.test(age1~Group, data=data_demos, na.action=na.omit)

```

# Sex
```{r}
# create table with sex x group informaion 
sex<-as.table(rbind(c(11,8), c(14,23), c(13,22))) #con M, con F, LTLE M, LTLE F, RTLE M, RTLE R

# chi square test 
chisq.test(sex) 
#the proportion is not sig different between the two groups 
#phi<-sqrt(chisq/n)

```

# Education 
## omnibus 
```{r}
# summary
data_demos %>%
  group_by(Group) %>%
  summarize(mean = mean(edu, na.rm = TRUE), sd = sd(edu, na.rm = TRUE), min = min(edu, na.rm = TRUE), max = max(edu, na.rm = TRUE), n = n())

# model 
model<-lm(edu~Group, data=data_demos, na.action=na.omit)

# test model assumptions 
plot(model) # residuals vs fitted should be straight, no bad outliers 
hist(residuals(model)) # residuals should be normal 
shapiro.test(residuals(model)) # residuals should be normal 
ncvTest(model) # test for heterogeneity of variance 

# anova table 
Anova(model, type="II")
```

## posthocs 
### LTLE vs RTLE 
```{r}

## t test assumptions 
## normality 
demos_TLE %>%
  group_by(Group) %>%
  shapiro_test(edu)
## variance
demos_TLE %>% levene_test(edu ~ Group)

#normality violated so use wilcoxon rank sum test
wilcox.test(edu~Group, data=demos_TLE, conf.int=TRUE)

```

### LTLE vs CON
```{r}
## subset the data 
demos_LTLE_CON<-subset(data_demos, Group == "LTLE" | Group == "CON")

## t test assumptions 
## normality 
demos_LTLE_CON %>%
  group_by(Group) %>%
  shapiro_test(edu)
## variance
demos_LTLE_CON %>% levene_test(edu ~ Group)

# normality violated, use nonparametric 
wilcox.test(edu~Group, data=demos_LTLE_CON, conf.int=TRUE)

```

### RTLE vs CON
```{r}

# subset 
demos_RTLE_CON<-subset(data_demos, Group == "RTLE" | Group == "CON")

## t test assumptions 
## normality 
demos_RTLE_CON %>%
  group_by(Group) %>%
  shapiro_test(edu)
## variance
demos_RTLE_CON %>% levene_test(edu ~ Group)

# t test 
t.test(edu~Group, data = demos_RTLE_CON, alternative = "two.sided", paired = FALSE, var.equal=TRUE)

```



# Age of onset
```{r}

# summary
demos_TLE %>%
  group_by(Group) %>%
  summarize(mean = mean(AOO, na.rm = TRUE), sd = sd(AOO, na.rm = TRUE), min = min(AOO, na.rm = TRUE), max = max(AOO, na.rm = TRUE), n = n())

## t test assumptions 
## normality 
demos_TLE %>%
  group_by(Group) %>%
  shapiro_test(AOO)
## variance
demos_TLE %>% levene_test(AOO ~ Group)

# nomrality violated, use nonparametric 
wilcox.test(AOO~Group, data=demos_TLE, conf.int=TRUE)

```

# Hemispheric Dominance 
```{r}
## create table of counts per group 
hemdom<-as.table(rbind(c(31,6), c(30,3))) # LTLE left hem, LTLE B, RTLE left hem, RTLE B. 2 RTLE hemdom is NA

#chisq.test(hemdom) ## marginal, note this may not be best test becuase low cell numbers 
fisher.test(hemdom) #use fisher test becuase a cell count is <5, can see there is a sig diff in cell counts 
## if CI includes number 1, not sig for odds ratio 

```

# Surgery Type
```{r}
# create table of counts per group 
surg<-as.table(rbind(c(34,3), c(30,5))) ## ATL LTLE, selective LTLE, ATL RTLE, selective RTLE
#chisq.test(surg)
fisher.test(surg) #use fisher test becuase a cell count is <5, can see no diff in surg type 
## if CI includes number 1, not sig for odds ratio 

```
# Presence of MTS
```{r}
# create table of counts per group 
MTS<-as.table(rbind(c(27,10), c(23,12))) ## MTS LTLE, no MTS LTLE, MTS RTLE, no MTS RTLE
chisq.test(MTS)

```

# Matrix Similarity Distribution 
```{r}

## t test assumptions 
## normality 
demos_TLE %>%
  group_by(Group) %>%
  shapiro_test(MatrixSim_64)
## variance
demos_TLE %>% levene_test(MatrixSim_64 ~ Group)

# t test 
t.test(MatrixSim_64~Group, data = demos_TLE, alternative = "two.sided", paired = FALSE) 

```

Distribution of material specific memory 
```{r}

## t test assumptions 
## normality 
demos_LTLE %>%
  shapiro_test(Pre_Verb)
demos_RTLE %>%
  shapiro_test(Pre_Vis)

## comparing material specific preop memory 
t.test(demos_LTLE$Pre_Verb, demos_RTLE$Pre_Vis, alternative = "two.sided", paired = FALSE) #ns


## t test assumptions 
## normality 
demos_LTLE %>%
  shapiro_test(VerbalFactor)
demos_RTLE %>%
  shapiro_test(VisualFactor)

## comparing material specific memory decline between LTLE and RTLE 
t.test(demos_LTLE$VerbalFactor, demos_RTLE$VisualFactor, alternative = "two.sided", paired = FALSE) #ns

```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

